---
title: "Plot1 for EDA PS 1"
author: "Harold Pollack"
date: "9/14/2018"
html_document: default
word_document: default
keep_md: true

---

```{r setup, include=FALSE}
library("dplyr")
library("knitr")
library("ggplot2")
library("lattice")
library("gridExtra")
library("ggpubr")

knitr::opts_chunk$set(echo = TRUE)
```

Read the UCI data. Then save the dates 2/1/2007 and 2/2/2007. Note that the date variable is dd/mm/y. So some care must be taken. ALso cache=True given size of the dataset.

```{r read data, cache=TRUE}
mydata <- read.csv("/Users/haroldpollack/Documents/coursera_EDA/pS1/household_power_consumption.txt",header=TRUE,sep=";")
mydata$date_formatted <-as.Date(mydata$Date,format="%d/%m/%Y")
str(mydata)
summary(mydata)
# table(mydata$Date)
#
#   now extract the two dates called for in the assignment
#
data_for_graphing <-subset(mydata,(Date== "1/2/2007" | Date == "2/2/2007"))
str(data_for_graphing)
summary(data_for_graphing)
# table(data_for_graphing$Date)
table(data_for_graphing$date_formatted)
```

make the initial histogram and save as plot1.png

```{R plots, cache=TRUE}
#
# note that data_for_graphing$Global_active_power is a factor. It has to be changed to numeric for graphing
#
# Global_active_power is a factor. So one must be careful converting it to numeric
#
Global_active_power_numeric <- as.numeric(as.character(data_for_graphing$Global_active_power)) 
hist(Global_active_power_numeric,col="red",xlab="Global Active Power (kilowatts)",main="Global Active Power",ylab="Frequency")
#
#   now save the png file
#
dev.copy(png,'/Users/haroldpollack/Documents/coursera_datascience3/Plot1/plot1.png')
dev.off()