---
title: "Plot4"
author: "Harold Pollack"
date: "9/15/2018"
output:
  word_document: default
  html_document: default
---

This is the markdown file that creates plot4. I didn't print out all the intermediate products that are the same as those for plot one.

```{r setup, include=FALSE}
library("dplyr")
library("knitr")
library("ggplot2")
library("lattice")
library("gridExtra")
library("ggpubr")

knitr::opts_chunk$set(echo = TRUE)
```

I first read the UCI data. Then save the dates 2/1/2007 and 2/2/2007. Note that the date variable is dd/mm/y. So some care must be taken. ALso cache=True given size of the dataset.

```{r read data, cache=TRUE}
mydata <- read.csv("/Users/haroldpollack/Documents/coursera_EDA/pS1/household_power_consumption.txt",header=TRUE,sep=";")
mydata$date_formatted <-as.Date(mydata$Date,format="%d/%m/%Y")
str(mydata)
# summary(mydata)
# table(mydata$Date)
#
#   now extract the two dates called for in the assignment
#
data_for_graphing <-subset(mydata,(Date== "1/2/2007" | Date == "2/2/2007"))
# str(data_for_graphing)
# summary(data_for_graphing)
# table(data_for_graphing$Date)
table(data_for_graphing$date_formatted)
#
# now figure out weekday
#
data_for_graphing$weekday <- weekdays(data_for_graphing$date_formatted)
table(data_for_graphing$weekday)
```

Now manipulate date/time. I used code found here http://biostat.mc.vanderbilt.edu/wiki/pub/Main/ColeBeck/datestimes.pdf 

```{R date-time manipulation, cache=TRUE}
data_for_graphing$dts1 <- paste(data_for_graphing$date_formatted, data_for_graphing$Time)
time_scr <-strptime(data_for_graphing$dts1,format = "%Y-%m-%d %H:%M:%S")
#
# time_scr uses strptime. I just included this to learn.
#
str(time_scr)
data_for_graphing$dts2 <- as.POSIXct(data_for_graphing$dts1, format = "%Y-%m-%d %H:%M:%S")
str(data_for_graphing$dts1)
str(data_for_graphing$dts2)
```

make the multi-facet graph and save as plot4.png
Note that I shrunk the legend using CEX

```{R plots, cache=TRUE}
#
# note that data_for_graphing$Global_active_power is a factor. It has to be changed to numeric for graphing
#
# Global_active_power is a factor. So one must be careful converting it to numeric
#
data_for_graphing$Global_active_power_numeric <- as.numeric(as.character(data_for_graphing$Global_active_power)) 
# hist(Global_active_power_numeric,col="red",xlab="Global Active  Power (kilowatts)",main="Global Active Power",ylab="Frequency")
data_for_graphing$Sub_metering_1_numeric <- as.numeric(as.character(data_for_graphing$Sub_metering_1)) 
summary(data_for_graphing$Sub_metering_1_numeric)
data_for_graphing$Sub_metering_2_numeric <- as.numeric(as.character(data_for_graphing$Sub_metering_2)) 
summary(data_for_graphing$Sub_metering_2_numeric)
summary(data_for_graphing$Sub_metering_3)

str(data_for_graphing$Sub_metering_1_numeric)
str(data_for_graphing$Sub_metering_2_numeric)
str(data_for_graphing$Sub_metering_3)

#
# time series by day
#
#
#    template here https://www.sixhat.net/plotting-multiple-data-series-in-r.html
#
par(mfrow=c(2,2))
with (data_for_graphing, {
  plot(dts2,as.numeric(as.character(Global_active_power)),type="l", xlab=" ", ylab = "Global activepower")     
   plot(dts2,as.numeric(as.character(Voltage)),type="l", xlab=" ", ylab = "voltage")
  plot(dts2,Sub_metering_1_numeric,type="l", xlab=" ", ylab = "Energy sub metering", col="blue")
points(dts2,Sub_metering_2_numeric, type='l',col="red")
points(dts2,Sub_metering_3, type='l',col="green")
legend ("topright",lty=1, col=c("blue","red","green"),legend = c("Sub_metering_1","Sub_metering_2","Sub_metering_3"),cex=0.5)
 
  plot(dts2,as.numeric(as.character(Global_reactive_power)),type="l", xlab=" ", ylab = "Energy sub metering")
       })



#
#   now save the png file
#
dev.copy(png,'/Users/haroldpollack/Documents/coursera_datascience3/Plot4/plot4.png')
dev.off()

